Machine Learning guided workflow for Ribosome Binding Site engineering

### 2021-05-28 paper/main.tex
Updates from Overleaf

Figure \ref{fig: Swarmplot and Quantplot}A shows the results for all the examined groups. 
In each round, we measure the TIR of benchmark RBS as the internal standard. 
We then obtain the normalised TIR (called \textit{TIR ratio}) by taking the ratio between the raw TIR and the average TIR of benchmark sequences in each round (which are run in triplicate in each round).
All Round 0 groups (BPS-NC, BPS-C, UNI, PPM, Bandit-0) have performed worse than our benchmark sequence in terms of TIR. 
The Bandit-0 group performed poorly, despite being machine learning driven, due to being trained on literature data.
However, starting from Round 1, where the bandit algorithm was fed data from the Round 0 the results become much better, with a number of sequences that perform better than the consensus Shine-Dalgarno sequence and in one case better than the benchmark (by 8\%).
In round 2 we have observed further improvement by getting more sequences that showed TIR on levels similar to our benchmark sequence.
Finally, in round 3 the algorithm identified two sequences that were 34\% and 15\% stronger than the benchmark sequence.
Figure \ref{fig: Swarmplot and Quantplot}B shows the same results but divided into quantiles where the specific point for a given group is showing the highest TIR for that quantile.
The gradual increase for all quantiles can be observed for all Bandit groups suggesting algorithms' better understanding of the experimental space with more data.
The decreased result in the 0.9th qunatile compared to the max value for Bandit 3 group can be attributed to the increased emphasis on exploitation that has been set for the that round compared to others.
We see that effect in Figure \ref{fig: Swarmplot and Quantplot} c), where we coloured the data points for Bandit 1-3 groups according to their relative exploration - exploitation affinity.
We can see the RBSs with high TIRs (blue, with high predicted mean) tend to come from exploitation the design space whereas the explorative points (red, with high predicted standard deviation) give relatively low TIR but expand our knowledge about the unknown part of the design space. 
Figure \ref{fig: Swarmplot and Quantplot}D shows the TIRs of RBSs tested in the Bandit groups divided into bins of width of 0.1 TIR ratio.
KDE plots have been overlaid to depict the calculated density for each group.
The increase of prevalence of later bandit groups in the higher bins is evident, especially for Bandit 2 and 3 constituting the bulk of results in the $>0.8$ TIR ratio bins.
Notably, the distributions calculated for all the groups show bivariate distribution, we discuss the possible reasons for that further in the text.\\
In \ref{fig: Swarmplot and Quantplot}E we show a tSNE plot depicting the experimental space.
Each RBS is located on the plot according to its distance to other RBSs as calculated by our kernel function.
We can see the RBSs recommended by Bandit groups have covered majority of the design space. 
A number of clusters were targeted by our recommendation algorithm.
For example, the circled clusters labelled as "G-Rich Clusters" have been actively recommended by the algorithm.
More specifically, sequences with more or equal to 4 guanines in any position constituted 10\% of the randomly selected sequences and 5, 9, 16 and finally 25\% in each of the 4 Bandit guided batches respectively.

@@

\begin{figure}[!ht]
    \centering
    \begin{subfigure}[b]{0.49\textwidth}
        \centering
        \caption{}
        \includegraphics[scale=0.4]{plots/Main_Paper/scatter_abc1_FF_0.pdf}
    \end{subfigure}
    % \hfill
    \begin{subfigure}[b]{0.49\textwidth}
        \centering
        \caption{}
        \includegraphics[scale=0.4]{plots/Main_Paper/scatter_abc1_FF_1.pdf}
    \end{subfigure}
    % \hfill
    \begin{subfigure}[b]{0.49\textwidth}
        \centering
        \caption{}
        \includegraphics[scale=0.4]{plots/Main_Paper/scatter_abc1_FF_2.pdf}
    \end{subfigure}
    \begin{subfigure}[b]{0.49\textwidth}
        \centering
        \caption{}
        \includegraphics[scale=0.4]{plots/Main_Paper/scatter_abc1_FF_3.pdf}
    \end{subfigure}
    \caption{\textbf{Predictions generalise to held out data.} The scatter plots A-D are showing the performance of our prediction algorithm (GPR with WDS kernel, with maximum k-mer as 6 and maximum shift as 1) calculated after each of the rounds.
    Note the TIR values are normalised TIR based on the standardisation described in section \ref{sec: method data pre-procesing}, which is different from the TIR ratio reported in the Figure \ref{fig: Swarmplot and Quantplot}.
    The values of 
    $R^2$ and Spearman correlation coefficient (with corresponding p-value) are all provided for each plot.
    The p-value decreases with round increases, which shows increasing strong evidence about the Spearman correlation between our prediction and measured TIR. 
    }
    \label{fig: Scatterplot}
\end{figure}

@@

\begin{table}[!h]
\centering
\begin{minipage}[c]{0.6\textwidth}
\centering
\begin{tabular}{|c|c|}
\hline
\textbf{Characteristics of the library}                                                       & \textbf{Statistics} \\ \hline
Total experimental space                                                                      & 4096                \\ \hline
Planned constructs                                                                            & 450                 \\ \hline
Successfully constructed                                                                      & 445                 \\ \hline
Sequences with CV\textless{}40\%                                                              & 79\%                \\ \hline
Sequences with CV\textless{}20\%                                                              & 27\%                \\ \hline
\begin{tabular}[c]{@{}c@{}}Efficiency of bandit design \\ (compared with random)\end{tabular} & 2                   \\ \hline
Raw TIR range                                                                                     &      [4.9258, 105.377]               \\ \hline
TIR ratio range                                                                                     &      [0.0626, 1.3393]               \\ \hline
\end{tabular}
\end{minipage}
\begin{minipage}[c]{0.38\textwidth}
\centering
% Please add the following required packages to your document preamble:
% \usepackage[table,xcdraw]{xcolor}
% If you use beamer only pass "xcolor=table" option, i.e. \documentclass[xcolor=table]{beamer}
\begin{tabular}{|c|c|}
\hline
\textbf{Top RBS Core} & \textbf{TIR Ratio} \\ \hline
GGGGGC                & 1.3393             \\ \hline
GGGGGT                & 1.1544             \\ \hline
GGCTAT                & 1.084              \\ \hline
\textbf{AGGAGA}                & 1                  \\ \hline
GGCGTT                & 0.9805             \\ \hline
GGGGGG                & 0.9788             \\ \hline
GGCGAC                & 0.9758             \\ \hline
CAGGAG                & 0.9629             \\ \hline
GGCGAG                & 0.9519             \\ \hline
\textbf{AGGAGG}                & 0.3941             \\ \hline
\end{tabular}
\end{minipage}
\caption{\textbf{Characteristics of the library.}
Left table presents some of the characteristics of our library.
Right table presents 10 RBS sequences with their corresponding TIR ratios, the first 9 being the strongest sequences including the benchmark sequence (AGGAGA) and the last being the Shine-Dalgarno sequence (AGGAGG).
CV is coefficient of variation (STD of a sample divided by its mean).
Efficiency of the bandit design is calculated by dividing the highest TIR found using machine learning by highest TIR found using random sequence generation. }
\end{table}

>>>>

Figure \ref{fig: Swarmplot and Quantplot}A shows the results for all the examined groups. 
In each round, we measure the TIR of benchmark RBS as the internal standard. 
We then obtain the normalised TIR (called \textit{TIR ratio}) by taking the ratio between the raw TIR and the average TIR of benchmark sequences in each round (which are run in triplicate in each round).
All Round 0 groups (BPS-NC, BPS-C, UNI, PPM, Bandit-0) have performed worse than our benchmark sequence in terms of TIR. 
The Bandit-0 group performed poorly, despite being machine learning driven, due to being trained on literature data.
However, starting from Round 1, where the bandit algorithm was fed data from the Round 0 the results become much better, with a number of sequences that perform better than the consensus Shine-Dalgarno sequence and in one case better than the benchmark (by 8\%).
In round 2 we have observed further improvement by getting more sequences that showed TIR on levels similar to our benchmark sequence.
Finally, in round 3 the algorithm identified two sequences that were 34\% and 15\% stronger than the benchmark sequence.
Figure \ref{fig: Swarmplot and Quantplot}B shows the same results but divided into quantiles where the specific point for a given group is showing the highest TIR for that quantile.
The gradual increase for all quantiles can be observed for all Bandit groups suggesting algorithms' better understanding of the experimental space with more data.
The decreased result in the 0.9th qunatile compared to the max value for Bandit 3 group can be attributed to the increased emphasis on exploitation that has been set for the that round compared to others.
We see that effect in Figure \ref{fig: Swarmplot and Quantplot} c), where we coloured the data points for Bandit 1-3 groups according to their relative exploration - exploitation affinity.
Those with high predicted mean are coloured blue and represents exploitation, those hued red are with high predicted uncertainty and represent exploration.
We can see the RBSs with high TIRs tend to come from exploitation the design space whereas the explorative points give relatively low TIR but expand our knowledge about the unknown part of the design space. 
% We can see the RBSs with high TIRs (blue, with high predicted mean) tend to come from exploitation the design space whereas the explorative points (red, with high predicted standard deviation) give relatively low TIR but expand our knowledge about the unknown part of the design space. 
Figure \ref{fig: Swarmplot and Quantplot}D shows the TIRs of RBSs tested in the Bandit groups divided into bins of width of 0.1 TIR ratio.
KDE plots have been overlaid to depict the calculated density for each group.
The increase of prevalence of later bandit groups in the higher bins is evident, especially for Bandit 2 and 3 constituting the bulk of results in the $>0.8$ TIR ratio bins.
Notably, the distributions calculated for all the groups show bivariate distribution, we discuss the possible reasons for that further in the text.\\
In \ref{fig: Swarmplot and Quantplot}E we show a tSNE plot depicting the experimental space.
Each RBS is located on the plot according to its distance to other RBSs as calculated by our kernel function.
We can see the RBSs recommended by Bandit groups have covered majority of the design space. 
A number of clusters were targeted by our recommendation algorithm.
For example, the circled clusters labelled as ``G-Rich Clusters" have been actively recommended by the algorithm.
More specifically, sequences with more or equal to 4 guanines in any position constituted 10\% of the randomly selected sequences and 5, 9, 16 and finally 25\% in each of the 4 Bandit guided batches respectively.

@@

\begin{figure}[!ht]
    \centering
    \begin{subfigure}[b]{0.49\textwidth}
        \centering
        \caption{}
        \includegraphics[scale=0.4]{plots/Main_Paper/scatter_abc1_FF_0.pdf}
    \end{subfigure}
    % \hfill
    \begin{subfigure}[b]{0.49\textwidth}
        \centering
        \caption{}
        \includegraphics[scale=0.4]{plots/Main_Paper/scatter_abc1_FF_1.pdf}
    \end{subfigure}
    % \hfill
    \begin{subfigure}[b]{0.49\textwidth}
        \centering
        \caption{}
        \includegraphics[scale=0.4]{plots/Main_Paper/scatter_abc1_FF_2.pdf}
    \end{subfigure}
    \begin{subfigure}[b]{0.49\textwidth}
        \centering
        \caption{}
        \includegraphics[scale=0.4]{plots/Main_Paper/scatter_abc1_FF_3.pdf}
    \end{subfigure}
    \caption{\textbf{Predictions generalise to held out data.} The scatter plots A-D are showing the performance of our prediction algorithm (GPR with WDS kernel, with maximum k-mer as 6 and maximum shift as 1) calculated after each of the rounds.
    Note the TIR values are normalised TIR based on the standardisation described in section \ref{sec: method data pre-procesing}, which is different from the TIR ratio reported in the Figure \ref{fig: Swarmplot and Quantplot}.
    The values of 
    $R^2$ and Spearman correlation coefficient (with corresponding p-value) are all provided for each plot.
    The p-value is for the hypothesis test whose null hypothesis is that two sets of data are uncorrelated.
    The p-value decreases with round increases, which shows increasing strong evidence about the Spearman correlation between our prediction and measured TIR. 
    }
    \label{fig: Scatterplot}
\end{figure}

@@

\begin{table}[!h]
\centering
\begin{minipage}[c]{0.6\textwidth}
\centering
\begin{tabular}{|c|c|}
\hline
\textbf{Characteristics of the library}                                                       & \textbf{Statistics} \\ \hline
Total experimental space                                                                      & 4096                \\ \hline
Planned constructs                                                                            & 450                 \\ \hline
Successfully constructed                                                                      & 445                 \\ \hline
Sequences with CV\textless{}40\%                                                              & 79\%                \\ \hline
Sequences with CV\textless{}20\%                                                              & 27\%                \\ \hline
\begin{tabular}[c]{@{}c@{}}Efficiency of bandit design \\ (compared with random)\end{tabular} & 2                   \\ \hline
Raw TIR range                                                                                     &      [4.926, 105.377]               \\ \hline
TIR ratio range                                                                                     &      [0.063, 1.339]               \\ \hline
\end{tabular}
\end{minipage}
\begin{minipage}[c]{0.38\textwidth}
\centering
% Please add the following required packages to your document preamble:
% \usepackage[table,xcdraw]{xcolor}
% If you use beamer only pass "xcolor=table" option, i.e. \documentclass[xcolor=table]{beamer}
\begin{tabular}{|c|c|}
\hline
\textbf{Top RBS Core} & \textbf{TIR Ratio} \\ \hline
GGGGGC                & 1.339             \\ \hline
GGGGGT                & 1.154             \\ \hline
GGCTAT                & 1.084              \\ \hline
\textbf{AGGAGA}                & 1                  \\ \hline
GGCGTT                & 0.981            \\ \hline
GGGGGG                & 0.979             \\ \hline
GGCGAC                & 0.976             \\ \hline
CAGGAG                & 0.963             \\ \hline
GGCGAG                & 0.952             \\ \hline
\textbf{AGGAGG}                & 0.394            \\ \hline
\end{tabular}
\end{minipage}
\caption{\textbf{Characteristics of the library.}
Left table presents some of the characteristics of our library.
Right table presents 10 RBS sequences with their corresponding TIR ratios, the first 9 being the strongest sequences including the benchmark sequence (AGGAGA) and the last being the Shine-Dalgarno sequence (AGGAGG).
CV is coefficient of variation (STD of a sample divided by its mean).
Efficiency of the bandit design is calculated by dividing the highest TIR found using machine learning by highest TIR found using random sequence generation. }
\end{table}
